
---
title: "Head to Head Shot Matrix"
output: html_document
---

```{r echo=FALSE, results='hide',message=FALSE}
library(reshape); library(stringr); library(lubridate); library(ggplot2);
library(rvest); library(XML); library(dplyr); library(plyr);
library(stringr); library(DataCombine); library(rvest); library(XML); library(dplyr); library(plyr);

d3.game <- function(season, game) {
  
  library(nhlscrapr);   library(d3heatmap)

  ###pull game from nhlscrapr
  select.game <- retrieve.game (season=season, gcode=game)
  
  game.title <- noquote(paste0(select.game[[3]][[1]]," ",select.game[[6]]["awayscore"]," @ ",
                       select.game[[3]][[2]]," ",select.game[[6]]["homescore"]," - ",
                       select.game[[4]][[2]]," ",select.game[[4]][[3]]," ",select.game[[4]][[4]]))

  ###limit to shot attempts
  processed.game <- select.game[[1]] %>%
            as.data.frame() %>%
            filter(etype == "GOAL" | etype == "SHOT" | etype == "MISS" | etype == "BLOCK") %>%
            filter(a5 != "" & h5 != "") %>%
          mutate(goal = ifelse(etype=="GOAL", 1, 0),
                   shot = 1) %>%
          select(a1, a2, a3, a4, a5, h1, h2, h3, h4, h5, ev.team, hometeam) 
 
  player.list <- select.game[[2]] %>%
      filter(pos != "G") %>%
      select(numfirstlast, hometeam)

  ###create player list for home & away
  home.player.list <- player.list %>% filter(hometeam == 1) %>% select(numfirstlast) %>% as.list()
  away.player.list <- player.list %>% filter(hometeam == 0) %>% select(numfirstlast) %>% as.list()

  home.player.list <- home.player.list[[1]]
  away.player.list <- away.player.list[[1]]
  
  ###create empty shot matrix
  shot.mat <- as.data.frame(matrix(ncol=length(away.player.list), nrow=length(home.player.list)))
  colnames(shot.mat) <- away.player.list
  rownames(shot.mat) <- home.player.list
  
  ###create shot matrix, for each home and away skater cumulative shot count for home player
  for(i in home.player.list) {

    for(j in away.player.list) {
  
             home.shot.tally.v1 <- processed.game %>%
                  filter((h1 == i | h2 == i | h3 == i | h4 == i | h5 == i) &
                           (a1 == j | a2 == j | a3 == j | a4 == j | a5 == j)) %>%
                  mutate(home.shot = ifelse(ev.team == hometeam, 1, -1))
          
             value <- ifelse(length(home.shot.tally.v1$home.shot) > 0,
                          sum(home.shot.tally.v1$home.shot),
                          NA)
             shot.mat[i, j] <- value
    }
  }
  
  ###ensure row & column has values
  shot.mat <- shot.mat[rowSums(is.na(shot.mat))!=ncol(shot.mat), ]
  shot.mat <- shot.mat[, colSums(is.na(shot.mat))!=nrow(shot.mat)]
  
  ###plot d3heatmap
  return(list(d3heatmap(shot.mat, xaxis_font_size = 8, yaxis_font_size = 8, na.rm = TRUE),game.title))
}

game.mat <- d3.game("20162017","20035")
```
Blue represents strong shot metrics for home team players (rows) relative to teammates.   Home player success ordered from top to bottom.   Away team players (columns) team relative shot attempt differentiate ordered left to right, red is preferable.

Dendrograms represent agglomerative hierarchical clustering is based on player shot attempt result similarity to teammates.
   Linemates are likely to be clustered together.

## `r noquote(game.mat[[2]])`

```{r echo=FALSE}
game.mat[[1]]
```
[Twitter](http://twitter.com/CrowdScoutSprts) @crowdscoutsprts

[Website: CrowdScout Sports](http://www.crowdscoutsports.com)